### 2020-07-31 code/figure1.Rmd

begin_NA <- data.frame(sample_id = paste0("sample", 1:50), 
                       clust_knn = min(as.numeric(mut_order$clust_knn)), ## put lowest cluster to prevent a new facet
                       log10mut = NA) ## put NA values so they don't get plotted
end_NA <- data.frame(sample_id = paste0("sample", 51:100), ## repeat the step for the ends
                       clust_knn = max(as.numeric(mut_order$clust_knn)),
                       log10mut = NA)
mut_order <- rbind(begin_NA, mut_order, end_NA) ## rbind the data in a specfic order

@@

<p style="FONT-SIZE:10px; COLOR:#000000; LINE-HEIGHT:12px;">**Figure 1:
Classification of tumors by pathway disruptions.** a) Each of 377 selected
Reactome pathways (columns) is classified as disrupted if one or more genes is
mutated in the tumor sample (rows). Tumors are grouped by tissue of origin using
standardized abrreviations from the TCGA project. b) MCA analysis was performed
restricted to the 36 informative components. c) The resulting MCA was projected
down to 3 dimensions for the purpose of density based clustering (only the first
2 dimensions are shown). Each dot represents a tumor d) HDBSCAN was used to
perform hierarchical density based clustering tuned to minimize the fraction of
unassigned tumors. e) HDBSCAN clustered tumors were visualized on the UMAP
projection; tumors are colored by cluster membership (see legend). This results
in some tumors remaining unclassified (grey points). f) kNN was used to assign
membership to the remaining tumors. g) Violin plots of sample mutation counts
for tumors in each class. h) Projection of cancer type identity onto UMAP.
Abbrevs: BRCA = Breast cancer, COAD = Colorectal adenocarcinoma, HNSC = Head and
Neck Squamous Cell, PAAD = Pancreatic adenocarcinoma. i) Projection of breast
cancer subtypes onto the UMAP. j) projection of BRCA1/2 somatic mutation onto the
UMAP. k) projection of tumor stage onto the UMAP,
regardless of cancer type. l) Projection of metastatic status onto the UMAP.
Abbrevs: M0 = non-metastatic tumors, M1 = metastatic tumors.</p>

>>>>

begin_NA <- data.frame(sample_id = paste0("sample", 1:100), 
                       clust_knn = min(as.numeric(mut_order$clust_knn)), ## put lowest cluster to prevent a new facet
                       log10mut = NA) ## put NA values so they don't get plotted
end_NA <- data.frame(sample_id = paste0("sample", 101:200), ## repeat the step for the ends
                       clust_knn = max(as.numeric(mut_order$clust_knn)),
                       log10mut = NA)
mut_order <- rbind(begin_NA, mut_order, end_NA) ## rbind the data in a specfic order

@@

<p style="FONT-SIZE:10px; COLOR:#000000; LINE-HEIGHT:12px;">**Figure 1:
Classification of tumors by pathway disruptions.** a) Each of 377 selected
Reactome pathways (columns) is classified as disrupted if one or more genes is
mutated in the tumor sample (rows). Tumors are grouped by tissue of origin using
standardized abbreviations from the TCGA project. b) MCA analysis was performed to 
determine the eigenvalue with the most explanatory power from the categorical data.
Average of 100 permutations of the data was used as a baseline. c) The p-value of
each eigenvalue was plotted and the max eigenvalue with a pvalue < 0.05 was selected. 
d) The resulting MCA was projected down to 3 dimensions for the purpose of density 
based clustering (only the first 2 dimensions are shown). Each dot represents a tumor 
e) HDBSCAN was used to perform hierarchical density based clustering tuned to minimize
the fraction of unassigned tumors. f) HDBSCAN clustered tumors were visualized on the UMAP
projection; tumors are colored by cluster membership (see legend). This results
in some tumors remaining unclassified (grey points). g) kNN was used to assign
membership to the remaining tumors. h) Mutational burden plot where each dot corresponds
to the sample mutation counts for tumors in each class. i) Projection of cancer 
type identity onto UMAP. Abbrevs: BRCA = Breast cancer, COAD = Colorectal adenocarcinoma,
HNSC = Head and neck squamous cell, PAAD = Pancreatic adenocarcinoma. j) Projection of breast
cancer subtypes onto the UMAP. k) Projection of BRCA1/2 somatic mutation onto the
UMAP. l) Projection of tumor stage onto the UMAP, regardless of cancer type. m) 
Projection of metastatic status onto the UMAP. Abbrevs: M0 = non-metastatic tumors, 
M1 = metastatic tumors.</p>
